---
date: "03/24/2022"
output:
  pdf_document:
    toc: TRUE
---


```{r, include = FALSE}
library(tidyverse)
library(GGally)
library(OOmisc)
library(MASS)
library(UsingR)
library(ggpubr)
library(rstatix)
library(tolerance)
```


# Problem 1 (Verzani problem 12.28)

*In the data set `mtcars` the variables `mpg`, `cyl`, and `am` indicate the
miles per gallon, the number of cylinders, and the type of transmission,
respectively. Perform a two-way ANOVA modeling `mpg` by the `cyl` and `am`, each
treated as a categorical variable.*

*Is there an indication of an interaction? Do both the number of cylinders and
the type of transmission make a difference?*

```{r, error = TRUE}

head(mtcars)
with(mtcars, interaction.plot(mpg, cyl, am))

carlcyl <- lm(mpg~as.factor(cyl), data = mtcars)
carlam <- lm(mpg~as.factor(am), data = mtcars)
carlcylam <- lm(mpg~am*as.factor(cyl), data = mtcars)

anova(carlam, carlcyl)
anova(carlcyl, carlcylam)
anova(carlam, carlcylam)
```
*These tests imply that there is no indication that transmission affects mpg in this dataset*
*Cylinders on the other hand do seem to have an affect on mpg*

# Problem 2 (Verzani problem 12.29)

*The data set `ToothGrowth` has measurements of tooth growth (`len`) of guinea
pigs for different dosages of Vitamin C (`dose`) and two different delivery
methods (`supp`).*

*Perform a two-way analysis of variance of tooth growth modeled by dosage and
delivery method. First, fit the full model including interactions and use the
$F$-test to compare this with the additive model.*

```{r, error = TRUE}
with(ToothGrowth, interaction.plot(dose, supp, len))

addit <- lm(len~supp+as.factor(dose), data = ToothGrowth)
inter <- lm(len~supp*as.factor(dose), data = ToothGrowth)

anova(addit, inter) #interaction has an effect

```

# Problem 3 (Verzani problem 12.30)

*The data set `OrchardSprays` contains measurements on the effectiveness of
various sprays on repelling honeybees. The variable `decrease` measures
effectiveness of the spray, `treatment` records the type of treatment, and
`rowpos` records the row in the field the measurement comes from.*

*Make an interaction plot of the mean of `decrease` with `treatment` as a trace
factor. Then fit the additive analysis-of-variance model and the model with
interaction. Compare the two models using `anova`. Is the interaction model
suggested by the results of the modeling?*

```{r, error = TRUE}
with(OrchardSprays, interaction.plot(rowpos, treatment, decrease))
interaction.plot()

additive <- lm(decrease~rowpos+treatment, data = OrchardSprays)
interaction <- lm(decrease~rowpos*treatment, data = OrchardSprays)

anova(additive, interaction)
#Giving odd NaN result, solution seems correct, so can't figure out why that is the case...
#Figured it out, apparently don't want them as.factor? Quirks!

```

# Problem 4 (Verzani problem 12.31)

*What does R output when there is not enough data to estimate the parameters and
an error term? We check with a simple example. In checking a comment that
everything is better with butter on it, a student asks four people to rate the
four combinations of bread and corn with and without butter. The data collected
is:*

```{r}
d <- data.frame(
  rating = c(8, 6, 8, 4),
  food = gl(2, 2, 4, labels = c("bread", "corn")),
  butter = gl(2, 1, 4, labels = c("yes", "no"))
)
xtabs(rating ~ butter + food, d)
with(d, interaction.plot(butter, food, rating))
```

*Look at the summary of the multiplicative model*

```{r}
summary(lm(rating ~ butter * food, d))
```

*Do the coefficients get estimated? What is the issue then?*

**YOUR COMMENTS HERE**

*Not enough data to populate a workable result with this amount of observations*
*It doesn't look like the lines really intersect on the graph either, apart from sharing 8*
*When there's not enough data to work with, the algorithm will populate null since there is a perfect*
*fit, resulting in no errors or a 0*
