---
title: "MATH 3080 Lab Project 3"
author: "Your Name Here"
date: "02/02/2022"
output:
  pdf_document:
    toc: TRUE
---

```{r, include=FALSE}
library(tidyverse)
library(GGally)
library(OOmisc)
library(MASS)
library(UsingR)
library(ggpubr)
library(rstatix)
```

# Problem 1 (Verzani problem 12.2)

*For the data set `Cars93` (**MASS**) perform a one-way analysis of variance of
`MPG.highway` for each level of `DriveTrain`. Does the data support the null
hypothesis of equal population means? (Use `aov()` for this problem.)*

```{r, error = TRUE}
#Take a quick look at data (Looks like we'll need to check assumptions)
boxplot(MPG.highway ~ DriveTrain, data = Cars93)

# Check assumption of normality (Shapiro's test states this isn't 
# normally distributed, but I'm assuming you want me to 
# move on anyway??)
model <- lm(MPG.highway ~ DriveTrain, data = Cars93)
ggqqplot(residuals(model))
shapiro_test(residuals(model))

# Check homogeneity of variances (pvalue is barely above an alpha 
# of 5%, so we can continue, though arguably with scrutiny (also, 
# still not normally distributed))
levene_test(MPG.highway ~ DriveTrain, data = Cars93)

# Perform ANOVA test (pvalue is quite low here, best to reject the null 
# (I'd assume anyway, tough to say with the outliers since not normally
# distributed))
res <- aov(MPG.highway ~ DriveTrain, data = Cars93)
summary(res)
```

# Problem 2 (Verzani problem 12.4)

*The data set `carsafety` (**UsingR**) contains car-crash data. For several
makes of cars the number of drivers killed per million is recorded in
`Drivers.deaths`. The number of drivers of other cars killed in accidents with
these cars, per million, is recorded in `Other.deaths`. The variable `type` is a
factor indicating the type of car.*

*Perform a one-way analysis of variance of the model `Drivers.deaths ~ type`. Is
there a difference in population means? Did you assume equal variances? Normally
distributed populations?*

*Repeat with an analysis of variance of the model `Other.deaths ~ type`. Is
there a difference in population means? (Use `oneway.test()` for this problem.)*

```{r, error = TRUE}
#Check assumption of normality for driver deaths (pass)
model2 <- lm(Driver.deaths ~ type, data = carsafety)
ggqqplot(residuals(model2))
shapiro_test(residuals(model2))

#Check homogeneity of variances for driver deaths (pass)
levene_test(Driver.deaths ~ type, data = carsafety)

#Perform ANOVA test (support null hypothesis)
oneway.test(Driver.deaths ~ type, data = carsafety)



#Check assumption of normality for other deaths (pass)
model3 <- lm(Other.deaths ~ type, data = carsafety)
ggqqplot(residuals(model3))
shapiro_test(residuals(model3))

#Check homogeneity of variances for other deaths (pass)
levene_test(Other.deaths ~ type, data = carsafety)

#Perform ANOVA test (support null hypothesis)
oneway.test(Other.deaths ~ type, data = carsafety)



```

# Problem 3 (Verzani problem 12.7)

*A manufacturer of point-of-sale merchandise tests three types of ENTER-button
markings. They wish to minimize wear, as customers get annoyed when the markings
on this button wear off. They construct a test of the three types, and conduct
several trials for each. The results, in unspecified units, are recorded in the
following table:*

|        |     |     |     |     |     |     |
|--------|-----|-----|-----|-----|-----|-----|
| Type 1 | 303 | 293 | 296 | 299 | 298 |     |
| Type 2 | 322 | 326 | 315 | 318 | 320 | 320 |
| Type 3 | 309 | 327 | 317 | 315 |     |     |

*Is there a difference in wear time among the three types? Answer this using a 
one-way ANOVA.*

```{r, error = TRUE}
# Your code here
buttonType <- c(1, 1, 1, 1, 1, 2, 2, 2, 2, 2,
                2, 3, 3, 3, 3)
wear <- c(303, 293, 296, 299, 298, 322, 326,
          315, 318, 320, 320, 309, 327, 317,
          315)

df <- data.frame(buttonType, wear)


# Check assumption of normality (pvalue is not greater than 5 
# - will not pass)
model4 <- lm(buttonType ~ wear, data = df)
ggqqplot(residuals(model4))
shapiro_test(residuals(model4))

# Dataset is also too small for a oneway.test
```

# Problem 4 (Verzani problem 12.13)

*For the data in Problem 2, perform the one-way ANOVA using `lm()`. Compare to
the results of `oneway.test()`.*

```{r, error = TRUE}
#Really interesting results from linear model tests 
# - they return pvalues far lower than oneway.test results

#Perform Linear Model test
summary(lm(Driver.deaths ~ type, data = carsafety))

#Perform Linear Model test
summary(lm(Other.deaths ~ type, data = carsafety))
```

# Problem 5 (Verzani problem 12.19)

*The `InsectSprays` data set contains a variable `count`, which counts the
number of insects and a factor `spray`, which indicates the treatment given.*

*First perform an analysis of variance to see whether the treatments make a
difference. If so, perform the Tukey HSD procedure to see which pairwise
treatments differ.*

```{r, error = TRUE}
res2 <- aov(count ~ spray, data = InsectSprays)
summary(res2)

plot(TukeyHSD(res2))
```

