---
date: "November 10, 2021"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

```{r}
set.seed(5292016)
```

## Problem 1 (Verzani problem 6.2)
*Roll a pair of dice. Let $X$ be the largest value shown on the two dice. Use `sample()` to simulate five values of $X$.*

```{r, echo=FALSE}
# This code block resets the random seed in order to get consistent results. This will be hidden in the final document, but YOU MUST NOT CHANGE THIS!
set.seed(5292016)
```

```{r, error=TRUE, tidy=TRUE}
#Create two dice
dice1 <- 1:6
dice2 <- 1:6

#Roll two dice
roll1 <- sample(rep(dice1, times = 1, 5))
roll2 <- sample(rep(dice2, times = 1, 5))

#print the value of each rolled dice
roll1
roll2

#find the highest die value in each roll
pmax(roll1, roll2)
```

## Problem 2 (Verzani problem 6.3)
*The National Basketball Association lottery to award the first pick in the draft is held by putting 1,000 balls into a hopper and selecting one. The teams with the worst records the previous year have a greater proportion of the balls. The data set `nba.draft` (**UsingR**) contains the ball allocation for the year 2002. Use `sample()` with Team as the data vector and `prob=Balls` to simulate the draft. What team do you select? Repeat until Golden State is chosen. How long did it take?*

```{r, echo=FALSE}
# This code block resets the random seed in order to get consistent results. This will be hidden in the final document, but YOU MUST NOT CHANGE THIS!
set.seed(5292016)
```

```{r, error=TRUE, tidy=TRUE}
# Load library
library(UsingR)

# Set starting counter to determine attempts until Golden State is chosen
count <- 0

# Repeat the draft until Golden State is chosen
repeat{
  pick <- sample(nba.draft$Team, 1, prob = nba.draft$Balls)
  count <- count + 1
  if (pick == "Golden State")
    break
}

# Number of iterations until Golden State was chosen
count


```

## Problem 3 (Verzani problem 6.23)
*Find the quintiles ($0^{\text{th}}$, $20^{\text{th}}$, $40^{\text{th}}$, $60^{\text{th}}$, $80^{\text{th}}$, and $100^{\text{th}}$ quantiles/percentiles) of the standard Normal distribution (this is a normal distribution with mean 0 and standard deviation 1).*

```{r, error=TRUE, tidy=TRUE}
# Place quintiles into a vector
QuintVec <- c(0, .20, .40, .60, .80, 1)

# Since normal we can use qnorm() to find distribution
qnorm(QuintVec, mean = 0, sd = 1)
```

## Problem 4
*Consider flipping a fair coin $n$ times and counting the number of time the coin lands heads-up. This is a binomial random variable, but it turns out that as $n$ gets large, this distribution can be approximated with a Normal distribution, where the mean is $0.5 \times n$ and the standard deviation is $0.5 \times \sqrt{n}$. We know this thanks to the central limit theorem. When flipping 3, 5, and 20 times, plot the pmf of the binomial random variable with the appropriate parameters (you may use my function `plot_pmf()` from the lecture notes). Superimpose on each of these plots the density curve of the Normal distribution with the appropriate parameters (use the `lines()` function to do so; an example for doing this is in the lecture notes). What do you notice? When does the approximation appear best?*

```{r, error=TRUE, tidy=TRUE}
plot_pmf <- function(q, p) {
    # This will plot a series of horizontal lines at q with height p, setting
    # the y limits to a reasonable heights
    plot(q, p, type = "h", xlab = "x", ylab = "probability", main = "pmf", ylim = c(0, 
        max(p) + 0.1))
    # Usually these plots have a dot at the end of the line; the point function
    # will add these dots to the plot created above
    points(q, p, pch = 16, cex = 2)
}

# Flipping when n = 3 times
plot_pmf(0:3, dbinom(0:3, size = 3, prob = 0.5))

# Flipping when n = 5 times
plot_pmf(0:5, dbinom(0:5, size = 5, prob = 0.5))

# Flipping when n = 20 times
plot_pmf(0:20, dbinom(0:20, size = 20, prob = 0.5))


# The approximation is best at 20, since it best fills the approximate curve

```



```{r, error=TRUE, tidy=TRUE}
# Your code here
```
