---
date: "October 21, 2021"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

## Problem 1 (Verzani problem 3.2)
*For the `michelson` (**MASS**) data set, produce a density plot comparing `Speed` between Experiments 1 and 2.*

```{r, tidy=TRUE, error=TRUE}
# Load michelson
library(MASS)
data(michelson)

#split speed and experiments
m<-split(michelson$Speed, michelson$Expt)

#pull experiment 1 and 2 as they relate to speed into separate vectors
speed1<-m$`1`
speed2<-m$`2`

#create density plot comparing the two
plot(density(speed1), ylim = c(0, .007), col = "blue")
lines(density(speed2), lty = 3, col = "red")


```

## Problem 2 (Verzani problem 3.4)
*Three students record the time spent on homework per class. Their data is:*

| Student| 1  | 2  | 3  | 4  | 5 |
|:-------|---:|---:|---:|---:|--:|
| Marsha | 25 | 0  | 45 | 90 | 0 |
| Bill   | 30 | 30 | 30 | 30 |   |
| Holly  | 15 | 0  | 90 | 0  |   |

*Use a list to store these values. Then create a boxplot to compare. (You must use `boxplot()`'s formula interface for this problem.)*

```{r, tidy=TRUE, error=TRUE}
##first vector in list is marsha score, then bill score, and so on
dF<-stack(list(Marsha = c(25,0,45,90,0), Bill = c(30,30,30,30), Holly = c(15,0,90,0)))

#Create box plot of students scores
boxplot(dF$values~dF$ind)


```

## Problem 3 (Verzani problem 3.5)
*A group of nursing students take turns measuring some basic assessments. Their data is:*

|          | Temp | Pulse | Systolic | Diastolic |
|----------|------|-------|----------|-----------|
| Jackie   | 98.2 | 96    |  134     | 90        |
| Florence | 98.6 | 56    |  120     | 80        |
| Mildred  | 98.2 | 76    |  150     | 95        |

*Create a data frame of these values. Will `plot()` and `boxplot()` produce the same graphic?*

```{r, tidy=TRUE, error=TRUE}
# Create all fields
students <- c("Jackie", "Florence", "Mildred")
temp <- c(98.2, 98.6, 98.2)
pulse <- c(96, 56, 76)
systolic <- c(134, 120, 150)
diastolic <- c(90, 80, 85)

# Create the dataframe
nurseGroup <- data.frame(students, temp, pulse, systolic, diastolic)

#No, plot and boxplot don't produce the same graphic
```

## Problem 4 (Verzani problem 3.8)
*The second argument to `split` can be a list of factors. The result is that all interactions (possible combinations) are used for the groups. In the `ToothGrowth` data set, growth (`len`) is measured for two types of supplements (`supp`) and three doses (`dose`). Split this `len` value into 6 groups.*

```{r, tidy=TRUE, error=TRUE}
# Import ToothGrowth
data("ToothGrowth")

#split to 6 groups
TGrow <- split(ToothGrowth$len, c(ToothGrowth$supp, ToothGrowth$dose))
```
