---
title: "MATH 3070 Lab Project 12"
author: "Lane Zaugg"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

*Remember: I expect to see commentary either in the text, in the code with comments created using `#`, or (preferably) both! **Failing to do so may result in lost points!***

## Problem 1 (Verzani problem 8.7)
*Of the last ten times you've dropped your toast, it landed sticky-side down nine times. If these are a random sample from the $\text{Ber}(p)$ distribution, find an 80% confidence interval for $p$, the probability of the stidy side landing down. (Use `binconf()` (**Hmisc**) to compute the score interval.)*

```{r, tidy = TRUE}
# Load library
library(Hmisc)

# Use binconf to solve at 80% confidence interval
binconf(9, 10, alpha = .2, method = "all")
```

## Problem 2 (Verzani problem 8.10)
*A survey is taken of 250 students, and a $\hat{p}$ of 0.45 is found. The same survey is repeated with 1000 students, and the same $\hat{p}$ is found. Compare the two 95% confidence intervals. What is the relationship? Is the margin of error for the second one four times smaller? If not, how much smaller is it? (Use `binom.test()` to answer this problem.)*

```{r, tidy = TRUE}
# Calculate binomial test for 250 students, p(hat) = .45, where x = 250*p
# Not quite exact, as binom won't allow me to enter a "half person" of 112.5, so I rounded x to 113.
binom.test(113, 250, p = .45, conf.level = .95)

# Calculate binomial test for 1000 students, p(hat) = .45, where x = 1000*p = 450
binom.test(450, 1000, p = .45, conf.level = .95)

## Confidence intervals overlap, no significant difference

## the Margin of error difference shows that it is twice ## as small, not four times.
moe1 <- 1.96 * sqrt(.45*(1-.45)) / sqrt(250)
moe2 <- 1.96 * sqrt(.45*(1-.45)) / sqrt(1000)
moe1
moe2
moe1 - moe2
```

## Problem 3 (Verzani problem 8.15)
*The `stud.recs` (**UsingR**) data set contains a sample of math SAT scores from some population in the variable `sat.m`. Find a 90% confidence interval for the mean math SAT score for this data. (Do not use `t.test(); find this confidence interval "by hand".)*

```{r, tidy = TRUE}
# import
library(UsingR)

# Assign data to variable
var <- stud.recs$sat.m

#Find upper and lower limits
upper <- mean(var) + 1.65 * (sd(var)/sqrt(length(var)))
lower <- mean(var) - 1.65 * (sd(var)/sqrt(length(var)))

# print interval
upper
lower

```

## Problem 4 (Verzani problem 8.15)
*For the `homedata` (**UsingR**) data set find 90% confidence intervals for both variables `y1970` and `y2000`, assuming the sample represents some population. Perform one sample t-test for each variable, use `t.test()`, but first discuss whether the model assumptions are appropriate (include some check of the assumptions, like a Q-Q plot).*

```{r, tidy = TRUE}
# Create local variables of data
y1970 <- homedata$y1970
y2000 <- homedata$y2000

# Check assumptions of 1970 and 2000
qqnorm(y1970)
qqline(y1970)

qqnorm(y2000)
qqline(y2000)

# Find 90% confidence intervals for each
t.test(y1970, conf.level = .90)
t.test(y2000, conf.level = .90)
```