---
title: "MATH 3070 Lab Project 5"
author: "Lane Zaugg"
date: "September 28, 2021"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

*Remember: I expect to see commentary either in the text, in the code with comments created using `#`, or (preferably) both! **Failing to do so may result in lost points!***


## Problem 1 (Verzani problem 4.7)
*Use the `subset()` function to return a data frame made from the `Cars93` (**MASS**) data frame consisting only of `non-USA` cars in origin, with 4 cylinders and a maximum price of \$15,000 or less.*

```{r, tidy=TRUE,error=TRUE}
# Import MASS and look at dataframe of Cars93
library(MASS)
data.frame(Cars93)

# Create subset of Cars93 dataframe that contains the specifics we want
noUsaDataF <- subset(Cars93, Cylinders == 4 & Origin == "non-USA" & Max.Price < 15)

# Print specific data.frame with the sought after results.
data.frame(noUsaDataF)

```

## Problem 2 (Verzani problem 4.10)
*Find the standard deviation of each variable in the `mtcars` data set.*

```{r, tidy=TRUE, error=TRUE}
# Looking at dataframe of mtcars
data.frame(mtcars)

# Calculating the standard deviation of each column
# of mtcars. (Columns are isolated using the '2')
apply(mtcars, 2, sd)

```

## Problem 3 (Verzani problem 4.11)
*Find the standard deviation for each numeric variable in `Cars93` (**MASS**).*

```{r, tidy=TRUE, error=TRUE}
# Looking at Cars93 again
data.frame(Cars93)

# Calculating the standard deviation of each column
# of mtcars. (Columns are isolated using the '2')
# Any non-numeric column values are represented with an "NA"
apply(Cars93, 2, sd)

```

## Problem 4
*You can find data describing various statistics on 2012 Olympic participating nations such as population and medal counts in the comma-separated value file (olympic-medals2012.csv). Load this dataset into R in a data frame and do the following:*

1. *Create a new data frame containing only the country's ISO code (a three-letter code for the country; USA is the United States of America, and DZA Algeria), name, population, and counts for gold, silver, and bronze. This data frame should be called `olympic2012`.*

```{r, tidy=TRUE, error=TRUE}
# Import CSV
medals <- read.csv("C:\\Users\\Mac\\Downloads\\olympic-medals2012.csv", header=TRUE, stringsAsFactors=FALSE)

# Look at medals dataframe
data.frame(medals)

# Isolating ISO, name, pop, and medal count to olympic2012
# by calling specific column numbers
dim(medals)
olympic2012 <- (medals[, c(1,3,4,10,11,12), ])

#print
data.frame(olympic2012)
```

2. *What proportion of countries earned no medals? Answer this using `olympic2012`.*

```{r, tidy=TRUE, error=TRUE}
# subset
subolympic2012 <- subset(olympic2012, Gold == 0 & Silver == 0 & Bronze == 0)

# print
data.frame(subolympic2012)

# look at number of rows to find number of countries that earned nothing (119)
dim(subolympic2012)

# divide 119 by total number of participants (204) in olympic2012
dim(olympic2012)

# 58.3%
119/204
```

