---
title: "MATH 3080 Lab Project 7"
author: "Lane Zaugg"
date: "02/24/2022"
output:
  pdf_document:
    toc: TRUE
---

```{r, include = FALSE}
library(tidyverse)
library(GGally)
library(OOmisc)
library(MASS)
library(UsingR)
library(ggpubr)
library(rstatix)
library(tolerance)

coef_conf_int <- function(fit, conf.level = 0.95) {
  beta <- coef(fit)
  se <- coef_stderr(fit)
  nu <- deg_free(fit)
  tstar <- qt((1 - conf.level) / 2, df = nu, lower.tail = FALSE)
  moe <- se * tstar
  cbind(lower = beta - moe, upper = beta + moe)
}

min_aic_model <- function(...) {
  models <- list(...)
  aic_list <- sapply(models, AIC)
  best_model <- models[[which.min(aic_list)]]
  best_model
}

```

# Problem 1 (Verzani problem 11.22)

*For the data set `trees`, model the `Volume` by the `Girth` and `Height`
variables. Does the model fit the data well?*

```{r, error = TRUE}
# Your code here
fit <- lm(Volume~Girth+Height, data = trees)
summary(fit)
#good fit because they are close to 1 r^2
```

# Problem 2 (Verzani problem 11.24)

*For the `deflection` (**UsingR**) data set, fit the quadratic model*

$$\text{Deflection} = \beta_0 + \beta_1 \text{Load} + \beta_2 \text{Load}^2 +
\epsilon.$$

*How well does this model fit the data? Compare to the linear model.*

```{r, error = TRUE}

fits <- lm(Deflection~Load, data = deflection)
fit2 <- lm(Deflection~Load+I(Load^2), data = deflection)

summary(fits)
summary(fit2)

#Model fits data well compared to its linear counterpart.
```

# Problem 3 (Verzani problem 11.25)

*The data set `kid.weights` contains age, weight, and height measurements for
several children. Fit the linear model*

$$\text{weight} = \beta_0 + \beta_1 \text{age} + \beta_2 \text{height} + \beta_3
\text{height}^2 + \beta_4 \text{height}^3 + \beta_5 \text{height}^4$$

*Use the partial $F$-test to select between this model and the nested models
found by using only first-, second-, and third-degree polynomials for `height`.*

```{r, error = TRUE}
fit3 <- lm(weight ~ age + height + I(height^2) + I(height^3) + I(height^4), data = kid.weights)
fitted3 <- lm(weight ~ age + height + I(height^2) + I(height^3), data = kid.weights)

#anova 
anova(fit3) 
#next model nested in anova to compare (make another model that only has first second third (not fourth))) Compare and see if significant.
anova(fitted3)

#They both look equally useful to me. 
```

# Problem 4 (Verzani problem 11.27)

*The data set `Cars93` (**MASS**) contains data on cars sold in the United
States in the year 1993. Fit a regression model with `MPG.city` modeled by the
numeric variables `EngineSize`, `Weight`, `Passengers`, and `Price`. Which
variables are marked as statistically significant by the marginal $t$-tests?
Which model is selected by the AIC?*

```{r, error = TRUE}
fit4 <- lm(MPG.city ~ EngineSize + Weight + Passengers + Price, data = Cars93)

summary(fit4)

coef_conf_int(fit4)
min_aic_model(fit4)

```

# Problem 5 (Verzani problem 11.29)

*The data set `baycheck` (**UsingR**) contains estimated populations for a
variety of Bay Checkerspot butterflies near California. A common model for
population dynamics is the Ricker model, for which $t$ is time in years:

$$N_{t + 1} = aN_t e^{bN_t} W_t,$$

*where $a$ and $b$ are parameters and $W_t$ is a lognormal multiplicative error.
This can be turned into a regression model by dividing by $N_t$ and then taking
logs of both sides to give*

$$\log\left(\frac{N_{t + 1}}{N_t}\right) = \log(a) + b N_t + \epsilon_t.$$

*Let $y_t$ be the left-hand side. This may be written as*

$$y_t = r\left(1 - \frac{N_t}{K}\right) + \epsilon_t,$$

*because $r$ can be interpreted as an unconstrained growth rate and $K$ as a
carrying capacity.*

*Fit the model to the `baycheck` data set and find values for $r$ and $K$. To
find $y_t$ you can do the following:*

```{r, error = TRUE}
d <- with(baycheck, {
  n <- length(year)
  yt <- log(Nt[-1]/Nt[-n])
  nt <- Nt[-n]
  data.frame(yt, nt)
})
```

*Recall that a negative index means all but that index.*

```{r, error = TRUE}
# Your code here
# simplified model to linear form. See intercept and slope in terms of r and k, intercept is r (estimate for r will be intercept). Slope is minus r over k. 

# yts aren't given, hence use the function.
```

