---
title: "MATH 3080 Lab Project 10"
author: "Lane Zaugg"
date: "04/06/2022"
output:
  pdf_document:
    toc: TRUE
---



```{r, include = FALSE}
library(BSDA)
library(tidyverse)
library(GGally)
library(OOmisc)
library(MASS)
library(UsingR)
library(ggpubr)
library(rstatix)
library(tolerance)
```

# Problem 1 (Verzani problem 9.17)

*The `exec.pay` (**UsingR**) data set contains data on the salaries of CEOs at
199 top companies in the United States. The amounts are in \$10,000s. The data
is not symmetric. Do a sign test to determine whether the median pay is more
than \$220,000.*

```{r, error = TRUE}

SIGN.test(x = exec.pay, md = 22, alternative = "greater")

```
**Reject the null, median pay is likely more than $220k**

# Problem 2 (Verzani problem 9.18)

*Repeat the previous exercise, using the signed-rank test on the log-transformed
data. Do you reach the same conclusion?*

```{r, error = TRUE}

logPay <- log10(exec.pay)

plot(density(logPay))
wilcox.test(logPay, mu = 22)

```
**From what I can tell, yes**

# Problem 3 (Verzani problem 9.20)

*If the sign test has fewer assumptions on the population, why wouldn't we
always use that instead of a $t$-test? The answer lies in the power of the sign
test to detect when the null hypothesis is false. The sign test will not reject
a false null as often as the $t$-test. The following commands will perform a
simulation comparing the two tests on data that has a $N(1, 2)$ distribution.
The significance tests performed are both*

$$H_0: \mu = 0$$

$$H_A: \mu > 0.$$

*Run the simulation. Is there a big difference between the two tests?*

```{r}
m <- 200; n <- 10

out <- replicate(m, {
  x <- rnorm(n, mean = 1, sd = 2)
  ttest <- t.test(x, mu = 0, alt = "greater")$p.value
  sgntest <- 1 - pbinom(sum(x > 0) - 1, n, 1/2)
  c(t.test = ifelse(ttest < 0.05, 1, 0),
    sign.test = ifelse(sgntest < 0.05, 1, 0))
})

res.t <- out["t.test",]
res.sign <- out["sign.test",]
results <- c(t = sum(res.t)/m, sign = sum(res.sign) / m)

results
```

**I'm assuming these are pVals? Neither are below a typical alpha of 0.05, however the sign test is considerably**
**lower than the t test in this case. I can see how the sign test would not reject a false null assuming it is usually lower than the t test. **

# Problem 4 (Verzani problem 9.31)

*For the `babies` (**UsingR**) data set, the variable `age` contains the
recorded mom's age and `dage` contains the dad's age for several different cases
in the sample. Do a significance test of the null hypothesis of equal ages
against a one-sided alternative that the dads are older in the sampled
population.*

```{r, error = TRUE}

t.test(babies$age, babies$dage, alternative = "less", paired = FALSE)

```
**Reject the null, Dads are likely older in this sample**

# Problem 5 (Verzani problem 9.32)

*The data set `normtemp` (**UsingR**) contains body measurements for 130
healthy, randomly selected individuals from some parent population. The variable
`temperature` contains normal body temperature data and the variable `gender`
contains gender information, with male coded as 1 and female as 2. Is the sample
difference across the two groups statistically significant?*

```{r, error = TRUE}
male <- normtemp %>%filter(gender == 1)
female <- normtemp %>%filter(gender == 2)
t.test(male$temperature, female$temperature)
```
**Difference is significant, reject the null**
