---
title: "MATH 3070 Lab Project 13"
author: "Lane Zaugg"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

*Remember: I expect to see commentary either in the text, in the code with comments created using `#`, or (preferably) both! **Failing to do so may result in lost points!***

## Problem 1 (Verzani problem 9.14)
*The data set `normtemp` (**UsingR**) contains measurements of 130 healthy, randomly selected individuals. The variable `temperature` contains normal body temperature. Does the data appear to come from a Normal distribution? If so, perform a $t$-test to see if the commonly assumed value of 98.6 degrees fahrenheit is correct. (Studies have suggested that 98.2 degrees fahrenheit is more accurate.)*

```{r, tidy = TRUE}
# Import
library(UsingR)

# Check sample is normal
qqnorm(normtemp$temperature)
qqline(normtemp$temperature)

# Perform t-test
t.test(normtemp$temperature)

#98.2 appears more accurate than the commonly assumed temperature
```

## Problem 2 (Verzani problem 9.4)
*In the United States in 2007, the proportion of adults age 21-24 who had no medical insurance was 28.1 percent. A survey of 75 recent college graduates in this age range finds that 40 percent are without insurance. Does this support a difference from the nationwide proportion? Perform a test of significance and report the p-value. Is it significant? (Perform this test "by hand", not using `prop.test()`.)*

```{r, tidy = TRUE}
#Solve for pValue by hand

#Variables
n = 75
pHat = .4
pNot = .281

# Have to see if we equal the null, or do not 
# equal null for the alternate hypothesis
z <- (pHat - pNot) / sqrt((pHat * (1-pHat))/ n)

#Find the p value
pval <- 2*pnorm(-abs(z))

pval

#At a 95% confidence interval (where alpha = .05), the pValue is less than alpha, rejecting the null hypothesis. 
```

## Problem 3 (Verzani problem 9.16)
*A one-sided, one-sample $t$-test will be performed. What sample size is needed to have a power of 0.80 for a significance level of 0.05 if `delta = 0.05` and the population standard deviation is assumed to be 5?*

```{r, tidy = TRUE}
# Run a power.t.test to find sample size
power.t.test(power = .8, delta = .05, sd = 5, sig.level = 0.05,
             alternative = "one.sided", type = "one.sample")

# Sample size of at least 61,826.93 is needed (or 61,827)
```

## Problem 4 (Verzani problem 9.31)
*For the `babies` (**UsingR**) data set, the variable `age` contains the recorded mom's age and `dage` contains the dad's age for several cases in the sample. Do a significance test of the null hypothesis of equal ages against a one-sided alternative that the dads are older in the sampled population.*

```{r, tidy = TRUE}
# Use t.test, whats the difference in the age and dage? Alternative is age of mom is less than age of dad, age of mom-age of # dad is greater or less than zero, so alternative should be less.

# Perform statistical test
t.test(babies$age, babies$dage, alternative = "less")

```