---
title: "MATH 3070 Lab Project 4"
author: "Lane Zaugg"
date: "September 16, 2021"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---
4
*Remember: I expect to see commentary either in the text, in the code with comments created using `#`, or (preferably) both! **Failing to do so may result in lost points!***

## Problem 1 (Verzani problem 2.43)
*The `time` variable in the `nym.2002` data set (**UsingR**) contains the time to finish the 2002 New York City Marathon for a random sample of the finishers.*

1. *What percent ran the race in under 3 hours? (note that the data is in minutes)*

```{r}
library(UsingR)
#your code here
data(nym.2002)

#assign all time values in nym.2002 to a vector
vec1<- (nym.2002$time)

#sort out all vector values greater than 180 (3 hours)
perc<- which(vec1 < 180)

#divide the number of values less than 180 by the total number of entries in the race (1000)
#multiply by 100 to get percentage result (2.6% in this case)
(length(perc)/1000) * 100
```

2. *What is the time cutoff for the top 10%? The top 25%?*

```{r}
# Notice that these are the 90th and 75th percentiles, respectively
# Your code here
vec2<- (nym.2002$time)

#find the 10th and 25th percentiles using quantile function to display those above the top 10% and top 25%
quantile(vec2, probs = c(.1, .25))

```

3. *What time cuts off the bottom 10%?*

```{r}
# Your code here
# find the 90th percentile, this will display the time, the greater of which are in the bottom 10%.
quantile(vec2, probs = .9)
```

## Problem 2 (Verzani problem 4.1)
*The data set `UScereal` (**MASS**) contains data on cereals sold in the United States in 1993. For this data set, answer the following questions using R (i.e. MUST answer questions using a code. Do not count by hand):*

1. *How many rows does the data frame have? Columns?*

```{r}
# Your code here
data(UScereal)

# assign UScereal to a matrix data type
df1<- data.frame(UScereal)

# function to find number of rows this data frame has
nrow(df1)

# function to find number of columns in the data frame
ncol(df1)

```

2. *How many different manufacturers are included?*

```{r}
# Your code here

?UScereal

# Assign mfr to vec3
vec3<- (UScereal$mfr)

# Remove duplicate mfr references in vec3
vec3_uninque<- unique(vec3)

# Length of vec3_unique gives us total unique mfrs
length(vec3_uninque)

```

3. *How many vitamin categories are included?*

```{r}
# Your code here
vec4<- (UScereal$vitamins)

#remove duplicates
vec4_unique <- unique(vec4)

#total of unique categories
length(vec4_unique)

```

4. *How many cereals have a sugar level above 10?*

```{r}
# Your code here
vec5<-(UScereal$sugars)

#filter out sugars below 10
filtered<- which(vec5 > 10)

#number of cereals above 10 sugar (too many!)
length(filtered)
```

5. *What is the mean calorie value for cereals with more than 5 grams of fat? Less than or equal to 5?*

```{r}
# Your code here
vec6<-(UScereal$fat)

#filter to greater than 5 fat
bigfat_filtered<- which(vec6 > 5)
#filter less than or equal to 5 fat
littlefat_filtered<-which(vec6 <= 5)

#get indexes 18 and 32 where fat is greater than 5
bigfat_filtered

#get all other indexes where fat is less or eq to 5
littlefat_filtered

cal_vec<-(UScereal$calories)

#use index of cal_vec to find mean of calories greater than 5
mean(c(cal_vec[18], cal_vec[32]))

#remove greater than 5 from vector
indices <- c(18,32)
lessthan<- cal_vec[-indices]

#mean of less than or equal to 5
mean(lessthan)


```

6. *What is the mean calorie value for cereals on the middle shelf (2)?*

```{r}
# Your code here

#get index of shelf 2 to shelfv vector
shelfv<- which(UScereal$shelf == 2)

#calorie vector
cal_vec<-(UScereal$calories)

#get only shelf 2 indexes for mean function
indices <- c(shelfv)
lessthan<- cal_vec[+indices]

#find mean of calorie value on middle shelf
mean(lessthan)

```

## Problem 3
*Create a data frame containing the data in the following table:*

| First    | Last      | Age |
|----------|-----------|-----|
| Marcus   | Holstein  | 23  |
| Samuel   | Adams     | 56  |
| Gus      | McPherson | 43  |
| Margaret | Olsen     | 41  |
| Zim      | Newbold   | 95  |

```{r}
# Your code here

#inputing into dataframe
datframe<-data.frame(First = c("Marcus","Samuel","Gus", "Margaret", "Zim"), Last = c("Holstein","Adams","McPherson", "Olsen", "Newbold"), Age = c(23,56,43,41,95))

datframe

```
