---
title: "MATH 3070 Lab Project 2"
author: "Lane Zaugg"
date: "September 1, 2021"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

*Remember: I expect to see commentary either in the text, in the code with comments created using `#`, or (preferably) both! **Failing to do so may result in lost points!***

## Problem 1 (Verzani problem 1.7)
*The `rivers` (**UsingR**) or any other data set is available after loading the package **UsingR**. Load the package, and inspect the data set. Scan the values to find the largest one.*

```{r}
# Your solution here
#load
library(UsingR)
data(rivers)

#return element with the largest value
max(rivers)

```

## Problem 2 (Verzani problem 2.4)
*Create the following sequences, using `:`, `seq()`, or `rep()` as appropriate:*

1. `"a" "a" "a" "a" "a" "a"`

```{r}
# Your solution here
rep("a",6)
```

2. `1 3 ... 99` *(the odd numbers)*

```{r}
# Your solution here
# by 2 to return odd-only values
seq(1,99, by = 2)

```

3. `1 1 1 2 2 2 3 3 3`

```{r}
# Your solution here

#collection of rep functions
c(rep(1,3),rep(2,3),rep(3,3))

```

## Problem 3 (Verzani problem 2.3)
*Let our small data set be* `2 5 4 10 8`.

1. *Enter this data into a data vector `x`.*

```{r}
# Your code here
x<-c(2,5,4,10,8)

```

2. *Find the square of each number.*

```{r}
# Your code here
# Square entire set
x^2

```

3. *Subtract 6 from each number.*

```{r}
# Your code here
x-6
```

4. *Subtract 9 from each number and then square the answer.*

```{r}
# Your code here
# Ensure squaring the result
c((x-9)^2)

```